---
title: "Staten Island Housing Price Prediction"
author: "Yizheng Li, Fuming Xu, Kunyu Liang, Tong Pang"
date: "2026-04-21"
output:
  html_document:
    df_print: paged
  pdf_document: default
subtitle: NYC Property Sales ML Project
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

## 1. Load Data and Preprocessing

This section loads the Staten Island rolling sales file, cleans the data, keeps selected residential property types, creates the `AGE` and `SALE.MONTH` features, and one-hot encodes categorical variables.

Main cleaning steps:

- Remove non-useful columns such as address, apartment number, lot, block, and present tax/building class fields.
- Remove sales below $50,000 (non-market transfers and intra-family deed transfers).
- Remove invalid rows with missing or zero values for year built, land square feet, and gross square feet.
- Keep selected residential building class categories.
- Add `AGE = sale year - year built`.
- One-hot encode ZIP code, tax class, sale month, neighborhood, building class category, and building class at time of sale.

```{r load_data}
# Packages
needed_packages <- c("readxl", "readr", "lubridate", "randomForest")
installed_packages <- rownames(installed.packages())

for (pkg in needed_packages) {
  if (!(pkg %in% installed_packages)) {
    install.packages(pkg, repos = "https://cloud.r-project.org")
  }
}

library(readxl)
library(readr)
library(lubridate)
library(randomForest)

# ---------------------------------------------------------
# Read data
# ---------------------------------------------------------

possible_excel_files <- list.files(
  pattern = "rollingsales_statenisland.*\\.xlsx$",
  ignore.case = TRUE
)

possible_csv_files <- list.files(
  pattern = "rollingsales_statenisland.*\\.csv$",
  ignore.case = TRUE
)

if (length(possible_excel_files) > 0) {
  data_file <- possible_excel_files[1]
  df <- read_excel(data_file, skip = 4)
  df <- as.data.frame(df)
} else if (length(possible_csv_files) > 0) {
  data_file <- possible_csv_files[1]
  df <- read_csv(data_file, skip = 4, show_col_types = FALSE)
  df <- as.data.frame(df)
} else {
  stop("Data file not found. Put rollingsales_statenisland Excel or CSV file in the same folder as this Rmd file.")
}

cat("Using file:", data_file, "\n")

names(df) <- make.names(trimws(names(df)))

cat("Raw rows:", nrow(df), "\n")
cat("Raw columns:", ncol(df), "\n")

# ---------------------------------------------------------
# Clean text columns
# ---------------------------------------------------------

text_cols <- c(
  "BUILDING.CLASS.CATEGORY",
  "BUILDING.CLASS.AT.TIME.OF.SALE",
  "NEIGHBORHOOD"
)

for (col in text_cols) {
  if (col %in% names(df)) {
    df[[col]] <- trimws(as.character(df[[col]]))
  }
}

# ---------------------------------------------------------
# Clean numeric columns
# ---------------------------------------------------------

num_cols_to_clean <- c(
  "SALE.PRICE",
  "GROSS.SQUARE.FEET",
  "LAND.SQUARE.FEET",
  "RESIDENTIAL.UNITS",
  "COMMERCIAL.UNITS",
  "YEAR.BUILT",
  "ZIP.CODE",
  "TAX.CLASS.AT.TIME.OF.SALE"
)

for (col in num_cols_to_clean) {
  if (col %in% names(df)) {
    df[[col]] <- parse_number(as.character(df[[col]]))
  }
}

# ---------------------------------------------------------
# Keep selected residential building class categories
# ---------------------------------------------------------

keep_categories <- c(
  "01 ONE FAMILY DWELLINGS",
  "02 TWO FAMILY DWELLINGS",
  "03 THREE FAMILY DWELLINGS",
  "04 TAX CLASS 1 CONDOS",
  "07 RENTALS - WALKUP APARTMENTS",
  "09 COOPS - WALKUP APARTMENTS",
  "10 COOPS - ELEVATOR APARTMENTS",
  "12 CONDOS - WALKUP APARTMENTS",
  "13 CONDOS - ELEVATOR APARTMENTS",
  "15 CONDOS - 2-10 UNIT RESIDENTIAL"
)

df <- df[df$BUILDING.CLASS.CATEGORY %in% keep_categories, ]

# ---------------------------------------------------------
# Basic cleaning
# ---------------------------------------------------------

df <- df[
  !is.na(df$SALE.PRICE) &
    df$SALE.PRICE >= 50000,
]

df <- df[
  !is.na(df$YEAR.BUILT) &
    df$YEAR.BUILT > 0,
]

df <- df[
  !is.na(df$GROSS.SQUARE.FEET) &
    df$GROSS.SQUARE.FEET > 0,
]

df <- df[
  !is.na(df$LAND.SQUARE.FEET) &
    df$LAND.SQUARE.FEET > 0,
]

# ---------------------------------------------------------
# Date features
# ---------------------------------------------------------

if (!(inherits(df$SALE.DATE, "Date") ||
      inherits(df$SALE.DATE, "POSIXct") ||
      inherits(df$SALE.DATE, "POSIXt"))) {
  df$SALE.DATE <- parse_date_time(
    as.character(df$SALE.DATE),
    orders = c("ymd", "mdy", "dmy")
  )
}

df$AGE <- year(df$SALE.DATE) - df$YEAR.BUILT
df$SALE.MONTH <- month(df$SALE.DATE)

df <- df[
  !is.na(df$AGE) &
    df$AGE >= 0,
]

df$YEAR.BUILT <- NULL

# ---------------------------------------------------------
# Remove unnecessary columns
# ---------------------------------------------------------

remove_cols <- c(
  "BOROUGH",
  "LOT",
  "EASEMENT",
  "ADDRESS",
  "APARTMENT.NUMBER",
  "TOTAL.UNITS",
  "BLOCK",
  "SALE.DATE",
  "TAX.CLASS.AT.PRESENT",
  "BUILDING.CLASS.AT.PRESENT"
)

for (col in remove_cols) {
  if (col %in% names(df)) {
    df[[col]] <- NULL
  }
}

# ---------------------------------------------------------
# Remove rows with missing required columns
# ---------------------------------------------------------

required_cols <- c(
  "SALE.PRICE",
  "GROSS.SQUARE.FEET",
  "LAND.SQUARE.FEET",
  "RESIDENTIAL.UNITS",
  "COMMERCIAL.UNITS",
  "AGE",
  "ZIP.CODE",
  "TAX.CLASS.AT.TIME.OF.SALE",
  "SALE.MONTH",
  "NEIGHBORHOOD",
  "BUILDING.CLASS.CATEGORY",
  "BUILDING.CLASS.AT.TIME.OF.SALE"
)

df <- df[complete.cases(df[, required_cols]), ]

# ---------------------------------------------------------
# One-hot encoding with one reference category dropped
# ---------------------------------------------------------

one_hot_drop_reference <- function(col_name, data) {

  if (!(col_name %in% names(data))) {
    return(data.frame())
  }

  data[[col_name]] <- as.factor(data[[col_name]])

  if (length(unique(data[[col_name]])) < 2) {
    return(data.frame())
  }

  formula_text <- paste("~", col_name)
  encoded <- model.matrix(as.formula(formula_text), data = data)

  encoded <- encoded[, colnames(encoded) != "(Intercept)", drop = FALSE]
  encoded <- as.data.frame(encoded)
  names(encoded) <- make.names(names(encoded))

  return(encoded)
}

df$ZIP.CODE <- as.factor(df$ZIP.CODE)
df$TAX.CLASS.AT.TIME.OF.SALE <- as.factor(df$TAX.CLASS.AT.TIME.OF.SALE)
df$SALE.MONTH <- as.factor(df$SALE.MONTH)
df$NEIGHBORHOOD <- as.factor(df$NEIGHBORHOOD)
df$BUILDING.CLASS.CATEGORY <- as.factor(df$BUILDING.CLASS.CATEGORY)
df$BUILDING.CLASS.AT.TIME.OF.SALE <- as.factor(df$BUILDING.CLASS.AT.TIME.OF.SALE)

zip_encoded <- one_hot_drop_reference("ZIP.CODE", df)
tax_encoded <- one_hot_drop_reference("TAX.CLASS.AT.TIME.OF.SALE", df)
month_encoded <- one_hot_drop_reference("SALE.MONTH", df)
neighborhood_encoded <- one_hot_drop_reference("NEIGHBORHOOD", df)
building_category_encoded <- one_hot_drop_reference("BUILDING.CLASS.CATEGORY", df)
building_timeofsale_encoded <- one_hot_drop_reference("BUILDING.CLASS.AT.TIME.OF.SALE", df)

df <- cbind(
  df,
  zip_encoded,
  tax_encoded,
  month_encoded,
  neighborhood_encoded,
  building_category_encoded,
  building_timeofsale_encoded
)

df$ZIP.CODE <- NULL
df$TAX.CLASS.AT.TIME.OF.SALE <- NULL
df$SALE.MONTH <- NULL
df$NEIGHBORHOOD <- NULL
df$BUILDING.CLASS.CATEGORY <- NULL
df$BUILDING.CLASS.AT.TIME.OF.SALE <- NULL

# ---------------------------------------------------------
# Define feature groups
# ---------------------------------------------------------

numeric_cols <- c(
  "GROSS.SQUARE.FEET",
  "LAND.SQUARE.FEET",
  "RESIDENTIAL.UNITS",
  "COMMERCIAL.UNITS",
  "AGE"
)

zip_cols <- grep("^ZIP.CODE", names(df), value = TRUE)
tax_cols <- grep("^TAX.CLASS.AT.TIME.OF.SALE", names(df), value = TRUE)
month_cols <- grep("^SALE.MONTH", names(df), value = TRUE)
neighborhood_cols <- grep("^NEIGHBORHOOD", names(df), value = TRUE)
building_category_cols <- grep("^BUILDING.CLASS.CATEGORY", names(df), value = TRUE)
building_timeofsale_cols <- grep("^BUILDING.CLASS.AT.TIME.OF.SALE", names(df), value = TRUE)

all_features <- c(
  numeric_cols,
  zip_cols,
  tax_cols,
  month_cols,
  neighborhood_cols,
  building_category_cols,
  building_timeofsale_cols
)

model_df <- df[, c(all_features, "SALE.PRICE")]
model_df <- na.omit(model_df)

# ---------------------------------------------------------
# Log-transform SALE.PRICE
# Log transformation stabilizes variance and reduces the
# influence of very expensive properties on model fitting.
# Predictions will be in log-dollars and converted back
# using exp() for reporting in dollar terms.
# ---------------------------------------------------------
model_df$LOG.SALE.PRICE <- log(model_df$SALE.PRICE)
model_df$SALE.PRICE     <- NULL

cat("Final rows after cleaning:", nrow(model_df), "\n")
cat("Final columns:", ncol(model_df), "\n")
cat("Features:", length(all_features), "\n")
cat("ZIP dummy columns:", length(zip_cols), "\n")
cat("Neighborhood dummy columns:", length(neighborhood_cols), "\n")
cat("Building category dummy columns:", length(building_category_cols), "\n")
cat("Building at time of sale dummy columns:", length(building_timeofsale_cols), "\n")

head(model_df, 5)
```

## 2. Train / Development / Final Test Split

```{r split}
set.seed(1)

n <- nrow(model_df)

all_index <- sample(1:n)

train_size <- floor(0.60 * n)
dev_size <- floor(0.20 * n)

train_index <- all_index[1:train_size]

dev_index <- all_index[
  (train_size + 1):(train_size + dev_size)
]

test_index <- all_index[
  (train_size + dev_size + 1):n
]

DataFrameTrain <- model_df[train_index, ]
DataFrameDev <- model_df[dev_index, ]
DataFrameFinalTest <- model_df[test_index, ]

cat("Training rows:", nrow(DataFrameTrain), "\n")
cat("Development rows:", nrow(DataFrameDev), "\n")
cat("Final test rows:", nrow(DataFrameFinalTest), "\n")

```

## 3. Evaluation Function

```{r evaluation_function}
# All models are trained on LOG.SALE.PRICE.
# evaluate_model takes log-scale actual and predicted values,
# converts predictions back to dollars using exp(),
# then computes RMSE and MAE in dollars for interpretability.
evaluate_model <- function(actual_log, predicted_log) {

  # Convert back to dollar scale
  actual_dollars    <- exp(actual_log)
  predicted_dollars <- exp(predicted_log)

  # Metrics in dollar scale
  mse   <- mean((actual_dollars - predicted_dollars)^2)
  rmse  <- sqrt(mse)
  mae   <- mean(abs(actual_dollars - predicted_dollars))

  # R2 computed on log scale (standard practice)
  ss_res <- sum((actual_log - predicted_log)^2)
  ss_tot <- sum((actual_log - mean(actual_log))^2)
  r2     <- 1 - ss_res / ss_tot

  return(list(
    MSE  = mse,
    RMSE = rmse,
    MAE  = mae,
    R2   = r2
  ))
}
```

## 4. Model 1: Multiple Linear Regression

```{r linear_regression}
formula_lr <- as.formula(
  paste("LOG.SALE.PRICE ~", paste(all_features, collapse = " + "))
)

LR_fit <- lm(formula_lr, data = DataFrameTrain)

predict_train_lr <- predict(LR_fit, newdata = DataFrameTrain)
predict_dev_lr   <- predict(LR_fit, newdata = DataFrameDev)

lr_train_metrics <- evaluate_model(DataFrameTrain$LOG.SALE.PRICE, predict_train_lr)
lr_dev_metrics   <- evaluate_model(DataFrameDev$LOG.SALE.PRICE,   predict_dev_lr)

cat("Linear Regression Training RMSE: $", round(lr_train_metrics$RMSE), "\n")
cat("Linear Regression Dev     RMSE: $", round(lr_dev_metrics$RMSE),   "\n")
cat("Linear Regression Dev     MAE:  $", round(lr_dev_metrics$MAE),    "\n")
cat("Linear Regression Dev     R2:  ",   round(lr_dev_metrics$R2, 4),  "\n")

cat("First 10 coefficients:\n")
print(head(coef(LR_fit), 10))
```

## 4.1 Linear Regression Residual Plot on Development Set

```{r linear_regression_residual_plot}
# Convert back to dollars for plotting
actual_dollars_dev_lr <- exp(DataFrameDev$LOG.SALE.PRICE)
pred_dollars_dev_lr   <- exp(predict_dev_lr)

plot(
  pred_dollars_dev_lr,
  actual_dollars_dev_lr - pred_dollars_dev_lr,
  main = "Linear Regression Residuals on Development Set",
  xlab = "Predicted Sale Price ($)",
  ylab = "Residuals ($)",
  col = "steelblue",
  pch = 19,
  cex = 0.5
)
abline(h = 0, col = "red", lwd = 2)
```

## 4.2 Linear Regression: Actual vs Predicted on Development Set

```{r linear_actual_vs_predicted}
plot(
  actual_dollars_dev_lr,
  pred_dollars_dev_lr,
  main = "Linear Regression: Actual vs Predicted",
  xlab = "Actual Sale Price ($)",
  ylab = "Predicted Sale Price ($)",
  col = "steelblue",
  pch = 19,
  cex = 0.5
)
abline(a = 0, b = 1, col = "red", lwd = 2)
```

## 5. Model 2: Polynomial Regression

```{r polynomial_regression}
cat_terms <- paste(
  c(
    zip_cols, tax_cols, month_cols, neighborhood_cols,
    building_category_cols, building_timeofsale_cols
  ),
  collapse = " + "
)

poly_results  <- data.frame(
  Degree           = integer(),
  Training_RMSE    = numeric(),
  Development_RMSE = numeric(),
  Development_MAE  = numeric(),
  Development_R2   = numeric()
)

for (degree in 1:4) {
  poly_terms <- paste(
    paste0("poly(", numeric_cols, ", ", degree, ", raw=TRUE)"),
    collapse = " + "
  )
  formula_poly_i <- as.formula(
    paste("LOG.SALE.PRICE ~", poly_terms, "+", cat_terms)
  )
  fit_poly_i   <- lm(formula_poly_i, data = DataFrameTrain)
  pred_train_i <- predict(fit_poly_i, newdata = DataFrameTrain)
  pred_dev_i   <- predict(fit_poly_i, newdata = DataFrameDev)

  train_m <- evaluate_model(DataFrameTrain$LOG.SALE.PRICE, pred_train_i)
  dev_m   <- evaluate_model(DataFrameDev$LOG.SALE.PRICE,   pred_dev_i)

  poly_results <- rbind(poly_results, data.frame(
    Degree           = degree,
    Training_RMSE    = train_m$RMSE,
    Development_RMSE = dev_m$RMSE,
    Development_MAE  = dev_m$MAE,
    Development_R2   = dev_m$R2
  ))
}

best_poly_degree <- poly_results$Degree[which.min(poly_results$Development_RMSE)]
cat("Best polynomial degree based on development RMSE:", best_poly_degree, "\n")
print(poly_results)
```

## 6. Model 3: Random Forest

```{r random_forest}
set.seed(1)

rf_grid <- expand.grid(
  ntree = c(300, 500),
  mtry = c(
    min(5, length(all_features)),
    min(10, length(all_features)),
    min(15, length(all_features))
  ),
  maxnodes = c(30, 50),
  nodesize = c(10)
)

rf_grid <- unique(rf_grid)

rf_results <- data.frame(
  ntree = integer(),
  mtry = integer(),
  maxnodes = integer(),
  nodesize = integer(),
  Training_MSE = numeric(),
  Development_MSE = numeric(),
  Training_RMSE = numeric(),
  Development_RMSE = numeric(),
  Development_MAE = numeric(),
  Development_R2 = numeric()
)

rf_models <- list()

for (i in 1:nrow(rf_grid)) {

  rf_fit_i <- randomForest(
    LOG.SALE.PRICE ~ .,
    data = DataFrameTrain,
    ntree = rf_grid$ntree[i],
    mtry = rf_grid$mtry[i],
    maxnodes = rf_grid$maxnodes[i],
    nodesize = rf_grid$nodesize[i],
    importance = TRUE
  )

  pred_train_i <- predict(rf_fit_i, newdata = DataFrameTrain)
  pred_dev_i <- predict(rf_fit_i, newdata = DataFrameDev)

  train_metrics_i <- evaluate_model(
    DataFrameTrain$LOG.SALE.PRICE,
    pred_train_i
  )

  dev_metrics_i <- evaluate_model(
    DataFrameDev$LOG.SALE.PRICE,
    pred_dev_i
  )

  rf_results <- rbind(
    rf_results,
    data.frame(
      ntree = rf_grid$ntree[i],
      mtry = rf_grid$mtry[i],
      maxnodes = rf_grid$maxnodes[i],
      nodesize = rf_grid$nodesize[i],
      Training_MSE = train_metrics_i$MSE,
      Development_MSE = dev_metrics_i$MSE,
      Training_RMSE = train_metrics_i$RMSE,
      Development_RMSE = dev_metrics_i$RMSE,
      Development_MAE = dev_metrics_i$MAE,
      Development_R2 = dev_metrics_i$R2
    )
  )

  rf_models[[i]] <- rf_fit_i
}

print(rf_results)

best_rf_index <- which.min(rf_results$Development_RMSE)
rf_fit <- rf_models[[best_rf_index]]
best_rf_setting <- rf_results[best_rf_index, ]


```

## 6.1 Best Random Forest Evaluation on Development Set

```{r random_forest_eval}
predict_train_rf <- predict(rf_fit, newdata = DataFrameTrain)
predict_dev_rf   <- predict(rf_fit, newdata = DataFrameDev)

rf_train_metrics <- evaluate_model(DataFrameTrain$LOG.SALE.PRICE, predict_train_rf)
rf_dev_metrics   <- evaluate_model(DataFrameDev$LOG.SALE.PRICE,   predict_dev_rf)

cat("Random Forest Training RMSE: $", round(rf_train_metrics$RMSE), "\n")
cat("Random Forest Dev     RMSE: $", round(rf_dev_metrics$RMSE),   "\n")
cat("Random Forest Dev     MAE:  $", round(rf_dev_metrics$MAE),    "\n")
cat("Random Forest Dev     R2:  ",   round(rf_dev_metrics$R2, 4),  "\n")
```

## 6.2 Random Forest Residual Plot on Development Set

```{r random_forest_residual_plot}
actual_dollars_dev_rf <- exp(DataFrameDev$LOG.SALE.PRICE)
pred_dollars_dev_rf   <- exp(predict_dev_rf)

plot(
  pred_dollars_dev_rf,
  actual_dollars_dev_rf - pred_dollars_dev_rf,
  main = "Random Forest Residuals on Development Set",
  xlab = "Predicted Sale Price ($)",
  ylab = "Residuals ($)",
  col = "forestgreen", pch = 19, cex = 0.5
)
abline(h = 0, col = "red", lwd = 2)
```

## 6.3 Random Forest: Actual vs Predicted on Development Set

```{r random_forest_actual_vs_predicted}
plot(
  actual_dollars_dev_rf,
  pred_dollars_dev_rf,
  main = "Random Forest: Actual vs Predicted",
  xlab = "Actual Sale Price ($)",
  ylab = "Predicted Sale Price ($)",
  col = "forestgreen", pch = 19, cex = 0.5
)
abline(a = 0, b = 1, col = "red", lwd = 2)
```

## 7. Development Set Model Summary

```{r development_model_summary}
development_summary <- data.frame(
  Model = c(
    "Linear Regression",
    "Random Forest"
  ),
  Training_RMSE = c(
    lr_train_metrics$RMSE,
    rf_train_metrics$RMSE
  ),
  Development_RMSE = c(
    lr_dev_metrics$RMSE,
    rf_dev_metrics$RMSE
  ),
  Development_MAE = c(
    lr_dev_metrics$MAE,
    rf_dev_metrics$MAE
  ),
  Development_R2 = c(
    lr_dev_metrics$R2,
    rf_dev_metrics$R2
  )
)

print(development_summary)

best_model_name <- development_summary$Model[
  which.min(development_summary$Development_RMSE)
]

cat("Best model chosen by development RMSE:", best_model_name, "\n")
```

## 8. Final ERM / Final Test Evaluation for Best Model

```{r final_test_evaluation}
# Use the already-trained Random Forest (trained on 60% training set)
# Evaluate ONCE on the final 20% test set

final_pred_log <- predict(rf_fit, newdata = DataFrameFinalTest)

# Convert predictions back to dollars
final_pred <- exp(final_pred_log)

# Evaluate on dollar scale
final_metrics <- evaluate_model(
  DataFrameFinalTest$LOG.SALE.PRICE,
  final_pred_log
)

final_summary <- data.frame(
  Best_Model      = "Random Forest",
  Final_Test_RMSE = final_metrics$RMSE,
  Final_Test_MAE  = final_metrics$MAE,
  Final_Test_MSE  = final_metrics$MSE,
  Final_Test_R2   = final_metrics$R2
)

print(final_summary)
cat("Final RMSE: $", round(final_metrics$RMSE), "\n")
cat("Final MAE:  $", round(final_metrics$MAE),  "\n")
cat("Final R2:  ",   round(final_metrics$R2, 4), "\n")
```

## 8.1 Best Model Residual Plot on Final Test Set

```{r best_model_residual_plot}
actual_dollars_test <- exp(DataFrameFinalTest$LOG.SALE.PRICE)

plot(
  final_pred,
  actual_dollars_test - final_pred,
  main = paste("Best Model Residuals:", best_model_name),
  xlab = "Predicted Sale Price ($)",
  ylab = "Residuals ($)",
  col = "steelblue", pch = 19, cex = 0.5
)
abline(h = 0, col = "red", lwd = 2)
```

## 8.2 Best Model: Actual vs Predicted on Final Test Set

```{r best_model_actual_vs_predicted}
plot(
  actual_dollars_test,
  final_pred,
  main = "Best Model: Actual vs Predicted on Final Test Set",
  xlab = "Actual Sale Price ($)",
  ylab = "Predicted Sale Price ($)",
  col = "steelblue", pch = 19, cex = 0.5
)
abline(a = 0, b = 1, col = "red", lwd = 2)
```

## 9. Random Forest Feature Set Comparison: ZIP vs Neighborhood

After Random Forest was confirmed as the best model, we run it
with three different location feature sets on the development
set to determine which location representation works best.

```{r rf_feature_comparison}
set.seed(1)

rf_feature_sets <- list(
  "ZIP only" = c(
    numeric_cols, zip_cols, tax_cols, month_cols,
    building_category_cols, building_timeofsale_cols
  ),
  "Neighborhood only" = c(
    numeric_cols, neighborhood_cols, tax_cols, month_cols,
    building_category_cols, building_timeofsale_cols
  ),
  "ZIP + Neighborhood" = c(
    numeric_cols, zip_cols, neighborhood_cols, tax_cols, month_cols,
    building_category_cols, building_timeofsale_cols
  )
)

rf_feature_results <- data.frame(
  Feature_Set      = character(),
  Development_RMSE = numeric(),
  Development_MAE  = numeric(),
  Development_R2   = numeric()
)

for (set_name in names(rf_feature_sets)) {
  current_features <- rf_feature_sets[[set_name]]

  train_data_exp <- DataFrameTrain[, c(current_features, "LOG.SALE.PRICE")]
  dev_features   <- DataFrameDev[, current_features, drop = FALSE]

  rf_exp <- randomForest(
    LOG.SALE.PRICE ~ .,
    data      = train_data_exp,
    ntree     = 500,
    mtry      = min(15, length(current_features)),
    maxnodes  = 50,
    nodesize  = 10
  )

  pred_dev_exp <- predict(rf_exp, newdata = dev_features)
  metrics_exp  <- evaluate_model(DataFrameDev$LOG.SALE.PRICE, pred_dev_exp)

  rf_feature_results <- rbind(
    rf_feature_results,
    data.frame(
      Feature_Set      = set_name,
      Development_RMSE = metrics_exp$RMSE,
      Development_MAE  = metrics_exp$MAE,
      Development_R2   = metrics_exp$R2
    )
  )
}

print(rf_feature_results)
```

## 9.1 Performance by Price Range (Best Model on Final Test Set)

```{r performance_by_price_range}
bins   <- c(0, 500000, 1000000, 2000000, Inf)
labels <- c("Below $500K", "$500K-$1M", "$1M-$2M", "Above $2M")

price_range_results <- data.frame(
  Price_Range = character(),
  n           = integer(),
  RMSE        = numeric(),
  MAE         = numeric()
)

test_actual <- actual_dollars_test

for (i in seq_along(labels)) {
  mask <- test_actual >= bins[i] & test_actual < bins[i + 1]
  n    <- sum(mask)
  if (n < 5) next

  rmse   <- sqrt(mean((test_actual[mask] - final_pred[mask])^2))
  mae    <- mean(abs(test_actual[mask] - final_pred[mask]))

  price_range_results <- rbind(
    price_range_results,
    data.frame(
      Price_Range = labels[i],
      n = n, RMSE = rmse, MAE = mae
    )
  )
}

print(price_range_results)
```

## 9.2 Performance by ZIP Code (Best Model on Final Test Set)

```{r performance_by_zip}
zip_test <- DataFrameFinalTest[, zip_cols, drop = FALSE]

get_zip_label <- function(row) {
  which_one <- which(row == 1)
  if (length(which_one) == 0) return("Reference ZIP")
  return(names(row)[which_one[1]])
}

test_zip_labels <- apply(zip_test, 1, get_zip_label)

zip_results <- data.frame(
  ZIP  = character(),
  n    = integer(),
  RMSE = numeric(),
  MAE  = numeric()
)

for (z in sort(unique(test_zip_labels))) {
  mask <- test_zip_labels == z
  n    <- sum(mask)
  if (n < 5) next

  rmse <- sqrt(mean((test_actual[mask] - final_pred[mask])^2))
  mae  <- mean(abs(test_actual[mask] - final_pred[mask]))

  zip_results <- rbind(
    zip_results,
    data.frame(ZIP = z, n = n, RMSE = rmse, MAE = mae)
  )
}

zip_results <- zip_results[order(zip_results$RMSE), ]
print(zip_results)
```
